### 2024-09-10 Manuscript/manuscript.tex
Adding the references file

\documentclass[12pt,letterpaper]{article}
\usepackage{fullpage}
\usepackage[top=1.5cm, bottom=3.5cm, left=2.2cm, right=2.2cm]{geometry}
\usepackage{amsmath,amsthm,amsfonts,amssymb,amscd, esint}
\usepackage{lastpage}
\usepackage{enumerate}
\usepackage{fancyhdr}
\usepackage{mathrsfs}
\usepackage{graphicx}
\usepackage{listings}
\usepackage{hyperref}
\usepackage[english]{babel}
\usepackage[table,xcdraw]{xcolor}
\usepackage{natbib}
\usepackage{enumitem}
\usepackage{yfonts}
\usepackage{tikz}
\usepackage{wrapfig}
\usepackage[normalem]{ulem}
\usepackage{multicol}
\useunder{\uline}{\ul}{}

\newcommand{\HRule}[1]{\rule{\linewidth}{#1}}
\setcounter{tocdepth}{5}
\setcounter{secnumdepth}{5}

\setlength{\parindent}{0.0in}
\setlength{\parskip}{0.05in}

% Edit these as appropriate
\newcommand\course{}
\newcommand\subject{Final Degree Project}
\newcommand\documenttitle{AI application for azophotoswitches' optimization with pharmacological interest}
\newcommand\hwnumber{1} 
\newcommand\NetIDb{Universitat Autònoma de Barcelona}

\begin{document}
\title{\vspace{5cm} \normalsize \textsc{\NetIDb}
		\\ [2.0cm]
		\HRule{0.5pt} \\
		\LARGE \textbf{\uppercase{\documenttitle}}
		\HRule{2pt} \\ [0.5cm]
		\normalsize \textsc{Sergio Castañeiras Morales\\NIU:1598456\\
		}\normalsize \vspace*{5\baselineskip}}

\date{\today}

\author{\large \textsc{\subject}}

\begin{titlepage}
\clearpage\maketitle
\thispagestyle{empty}
\end{titlepage}

\newpage
\pagestyle{fancyplain}
\headheight 35pt
\lhead{\NetIDb}    
\rhead{\subject}
\cfoot{}
\rfoot{\small\thepage}
\headsep 1.5em
The current manuscript's format is temporal!!!\par
Testing cite: \cite{NationalLibraryOfMedicineGeneralArticle}

\newpage
%Bibliography %
\bibliographystyle{plain}
\bibliography{references} 
%%%%%%%%

\end{document}