### 2025-05-03 Manuscript/GeneralSources/decisionTree.tex
writting

\documentclass{standalone}
\usepackage[edges]{forest}
\usepackage{amsmath}
\usepackage{tikz}
\usetikzlibrary{shadows.blur,arrows.meta}

\begin{document}

\begin{forest}
%for tree={
%  grow=east,
%  draw,
%  rounded corners,
%  text width=4.5cm,
%  align=flush center,
%  parent anchor=east,
%  child anchor=west,
%  edge={->, thick},
%  l sep=10mm,
%  s sep=10mm
%}
[Does the molecule have no symmetry elements other than $E$?
  [Yes\\\textbf{Point group: } $C_1$]
  [No
    [Does the molecule have only a mirror plane?
      [Yes\\\textbf{Point group: } $C_s$]
      [No
        [Does the molecule have only an inversion center?
          [Yes\\\textbf{Point group: } $C_i$]
          [No
            [Does the molecule have a proper rotation axis $C_n$?
              [No\\\textbf{Unclassified: recheck symmetry elements}]
              [Yes
                [Does it have $n$ vertical mirror planes ($\sigma_v$)?
                  [Yes\\\textbf{Point group: } $C_{nv}$]
                  [No
                    [Does it have a horizontal mirror plane ($\sigma_h$)?
                      [Yes\\\textbf{Point group: } $C_{nh}$]
                      [No\\\textbf{Point group: } $C_n$]
                    ]
                  ]
                ]
              ]
            ]
          ]
        ]
      ]
    ]
  ]
]
\end{forest}

\end{document}